\documentclass[11pt, a4paper]{article}
%\usepackage{geometry}
\usepackage[inner=1.5cm,outer=1.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\pagestyle{empty}
\usepackage{graphicx}
\usepackage{fancyhdr, lastpage, bbding, pmboxdraw}
\usepackage[usenames,dvipsnames]{color}
\definecolor{darkblue}{rgb}{0,0,.6}
\definecolor{darkred}{rgb}{.7,0,0}
\definecolor{darkgreen}{rgb}{0,.6,0}
\definecolor{red}{rgb}{.98,0,0}
\usepackage[hyphens]{url}
\usepackage[breaklinks=true,colorlinks,pagebackref,pdfusetitle,urlcolor=darkblue,citecolor=darkblue,linkcolor=darkred,bookmarksnumbered,plainpages=false]{hyperref}
%\usepackage{breakurl}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\pagestyle{fancyplain}
\fancyhf{}
\lhead{ \fancyplain{}{Citizen FOSS: What Snowden Knew} }
%\chead{ \fancyplain{}{} }
\rhead{ \fancyplain{}{Yale ISP Reading Group, Spring 2017} }
%\rfoot{\fancyplain{}{page \thepage\ of \pageref{LastPage}}}
\fancyfoot[RO, LE] {page \thepage\ of \pageref{LastPage} }
\thispagestyle{plain}

%%%%%%%%%%%% LISTING %%%
\usepackage{listings}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
\usepackage{verbatim} % used to display code
\usepackage{fancyvrb}
\usepackage{acronym}
\usepackage{amsthm}
\VerbatimFootnotes % Required, otherwise verbatim does not work in footnotes!



\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{cmyk}{0.62,0.57,0.23,0}
\definecolor{lightlightgray}{gray}{0.93}



\lstset{
%language=bash,                          % Code langugage
basicstyle=\ttfamily,                   % Code font, Examples: \footnotesize, \ttfamily
keywordstyle=\color{OliveGreen},        % Keywords font ('*' = uppercase)
commentstyle=\color{gray},              % Comments font
numbers=left,                           % Line nums position
numberstyle=\tiny,                      % Line-numbers fonts
stepnumber=1,                           % Step between two line-numbers
numbersep=5pt,                          % How far are line-numbers from code
backgroundcolor=\color{lightlightgray}, % Choose background color
frame=none,                             % A frame around the code
tabsize=2,                              % Default tab size
captionpos=t,                           % Caption-position = bottom
breaklines=true,                        % Automatic line breaking?
breakatwhitespace=false,                % Automatic breaks only at whitespace?
showspaces=false,                       % Dont make spaces visible
showtabs=false,                         % Dont make tabls visible
columns=flexible,                       % Column format
morekeywords={__global__, __device__},  % CUDA specific keywords
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{center}
{\Large \textsc{Citizen FOSS: What Snowden Knew}}
\end{center}
\begin{center}
Spring 2017
\end{center}
%\date{December 8, 2016}

\begin{center}
\rule{6.5in}{0.4pt}
\begin{minipage}[t]{.75\textwidth}
\begin{tabular}{llcccll}
\textbf{Facilitator:} & Sean O'Brien & & &  & \textbf{Time:} & Wed. 5:00pm -- 7:00pm \\
\textbf{E-mail:} &  \href{mailto:sean.obrien@yale.edu}{sean.obrien@yale.edu} & & & & \textbf{Place:} & 40 Ashmun A436, New Haven, CT
\end{tabular}
\end{minipage}
\rule{6.5in}{0.4pt}
\end{center}
\vspace{.5cm}
\setlength{\unitlength}{1in}
\renewcommand{\arraystretch}{2}

\noindent\textbf{Statement:}  This reading group will explore Digital Security and Operations Security concepts within the context of pervasive corporate and government surveillance, a reality exposed most prominently by the 2013 Edward Snowden disclosures.  Snowden relied upon a mix of Free/Open-Source Software (FOSS) to communicate extremely sensitive data, despite powerful adversaries, because he simply ``couldn't trust'' the proprietary alternatives.  The contemporary lawyer and legal scholar may not be hunted daily by a nation-state but, nonetheless, faces a sea of complex software choices that effectively safeguard or undermine her civil liberties or those of her clients.  This digital maze is further complicated by the increasing frequency and escalation of cyber attacks, massive data breaches, and the threat of global cyberwar on the horizon.
\vskip.15in
\noindent\textbf{Objectives:}  We try to be as comprehensive as possible, discussing a wide variety of FOSS tools/applications with accompanying real-world examples.   
At the end of the course, a participant should be able to:
\begin{enumerate}
\item Understand the significance of privacy as it applies to software-mediated communication.
\item Describe the general scope of Five Eyes surveillance and that of its corporate partners.
\item Perceive the importance of FOSS, transparent development, and open technology as security principle.
\item Evaluate emerging technology on its merit and potential for data security, privacy, and anonymity.
\item Initiate and sustain encrypted communication, often over networks that safeguard anonymity.
\item Apply simple techniques to everyday Web browsing that improve user privacy and data security.
\item Explore the digital frontiers of GNU/Linux, P2P, Tor, and the Deep Web. 
\item Develop a communications plan for real-world implementation of privacy-respecting technology.
\end{enumerate}

\vskip.07in
\noindent\textbf{Schedule:} Wednesdays 5:00pm -- 7:00pm, roughly every other week (see sessions below for specific dates).

\vskip.25in
\noindent\textbf{Week 1 - Why Free/Open-Source Software Matters} (January 25, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Software Freedom Law Center, \emph{A Legal Issues Primer for Open Source and Free Software Projects} (2008), Chapters 1 \& 2:  \url{https://www.softwarefreedom.org/resources/2008/foss-primer.html}

    \item Edward Snowden, ``The Last Lighthouse: Free Software in Dark Times,'' Mar. 19, 2016: \newline
    \url{http://ioterror.com/items/show/34}
    
    \item Bruce Schneier and Eben Moglen, ``Snowden, the NSA, and Free Software,'' Dec. 12, 2013:
    \newline
    \url{http://ioterror.com/items/show/4}
\end{itemize}

\textbf{In-class Video:}
\vskip.04in
\indent Richard Stallman, ``Introduction to Free Software and the Liberation of Cyberspace,'' Apr. 7, 2014:
\newline \indent \url{https://www.fsf.org/blogs/rms/20140407-geneva-tedx-talk-free-software-free-society}
\vskip.15in
\textbf{Hands-on with:}
    Etherpad, Jitsi Meet, Tox

\vskip.15in
\noindent\textbf{Week 2 - Going Dark} (February 8, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Berkman Center for Internet \& Society, \emph{``Don't Panic''
Making Progress on the ``Going Dark'' Debate
} (2016), pp. 1-12:  \url{https://cyber.harvard.edu/pubrelease/dont-panic/Dont_Panic_Making_Progress_on_Going_Dark_Debate.pdf}

    \item Asher Schechter, ``Who Needs the KGB when we have Facebook?
An Interview with Eben Moglen,'' Apr. 8, 2015: \url{http://emoglen.law.columbia.edu/my_pubs/Who-needs-KGB-when-we-have-Facebook-Schechter.pdf}
    
    \item National Security Agency, ``Tor Stinks,'' Jun. 2012: 
    \newline
    \url{https://www.eff.org/files/2014/04/09/20131004-guard-tor_stinks.pdf}
\end{itemize}

\textbf{In-class Video:}
\newline
\indent Tor Project, ``Introduction to Tor,'' Mar. 17, 2015:  \url{https://www.youtube.com/watch?v=JWII85UlzKw}
\vskip.15in
\textbf{Hands-on with:}
    Better Web browsing \& search, Ad-blockers, Tor Browser Bundle

\vskip.25in
\noindent\textbf{Week 3 - Operating System Insecurity} (February 22, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Tom Mendelsohn, ``Secure Boot snafu: Microsoft leaks backdoor key,'' Aug. 11. 2016: 
    \newline
    \url{http://arstechnica.com/security/2016/08/microsoft-secure-boot-firmware-snafu-leaks-golden-key/}

    \item Daniel Kahn Gillmor, ``Is This the FBI's `New' Method for Unlocking the San Bernardino iPhone?,'' Mar. 22, 2016: \url{https://www.aclu.org/blog/free-future/fbis-new-method-unlocking-san-bernardino-iphone}
    
    \item Jacob Appelbaum, ``To Protect and Infect, the Militarization of the Internet,'' Dec. 31, 2013: 
    \newline
    \url{http://www.nakedcapitalism.com/2014/01/jacob-appelbaum-30c3-protect-infect-militarization-internet-transcript.html}
\end{itemize}

\textbf{In-class Video:}
\newline
\indent The Linux Gamer, ``What Is Linux?,'' Sep. 22, 2015: \url{https://www.youtube.com/watch?v=tFFNiMV27VY}
\vskip.15in
\textbf{Hands-on with:}
    Tails

\vskip.25in
\noindent\textbf{Week 4 - The Spy In Your Pocket} (March 8, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Ron Amadeo, ``Google's iron grip on Android: Controlling open source by any means necessary,'' Oct. 20, 2013:  \url{http://arstechnica.com/gadgets/2013/10/googles-iron-grip-on-android-controlling-open-source-by-any-means-necessary/4/}

    \item E-mail from Lisa Jackson to John Podesta, \emph{Wikileaks Podesta E-mails Archive}, Dec. 20, 2015: 
    \newline
    \url{https://wikileaks.org/podesta-emails/emailid/30593#efmAHtANd}
    
    \item Tobias Boelter, ``WhatsApp vulnerability explained: by the man who discovered it,'' Jan. 16, 2017: 
    \url{https://www.theguardian.com/technology/2017/jan/16/whatsapp-vulnerability-facebook}
\end{itemize}

\textbf{In-class Video:}
\newline
\indent Jonas Anton \"Ostman, ``Liberating Software at the Lower Levels'' (first 11 mins), Feb. 8, 2016: 
\newline
\indent \url{https://www.youtube.com/watch?v=1ZIX1zO7pAs}
\vskip.15in
\textbf{Hands-on with:}
    F-Droid, Orfox, GNU Ring

\vskip.15in
\noindent\textbf{Week 5 - Sharing On A Hostile Web} (March 22, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Barton Gellman and Ashkan Soltani, ``NSA infiltrates links to Yahoo, Google data centers worldwide,'' Oct. 30, 2013: \url{https://www.washingtonpost.com/world/national-security/nsa-infiltrates-links-to-yahoo-google-data-centers-worldwide-snowden-documents-say/2013/10/30/e51d661e-4166-11e3-8b74-d89d714ca4dd_story.html}

    \item Samuel Gibbs, ``Dropbox hack leads to leaking of 68m user passwords on the internet,'' Aug. 31, 2016: 
    \newline
    \url{https://www.theguardian.com/technology/2016/aug/31/dropbox-hack-passwords-68m-data-breach}
    
    \item Electronic Frontier Foundation, \emph{Encrypt The Web Report}, Nov. 4, 2014: 
    \newline
    \url{https://www.eff.org/encrypt-the-web-report}
\end{itemize}

\textbf{In-class Video:}
\newline
\indent Code.org, ``The Internet: Encryption \& Public Keys,'' Aug. 21, 2015: 
    \newline
\indent \url{https://www.youtube.com/watch?v=ZghMPWGXexs}
\vskip.15in
\textbf{Hands-on with:}
    Up1, OnionShare, SparkleShare

\vskip.25in
\noindent\textbf{Week 6 - Plugging The E-mail Hole} (April 5, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Arun Vishwanath, ```Spearphishing' Roiled the Presidential Campaign,'' Nov. 8, 2016:  \url{https://theconversation.com/spearphishing-roiled-the-presidential-campaign-heres-how-to-protect-yourself-68274}
    
    \item Matthias Pfau, ``Why a Private Key Should Not Be Stored on a Central Server,'' 
    Jan. 25, 2016: 
    \newline
    \url{https://tutanota.com/blog/posts/private-key}
    
    \item H. Abelson et al., \emph{The Risks of Key Recovery, Key Escrow, and Trusted Third-Party Encryption (Revised)} (1998), Chapters 1\& 2:  \url{https://www.schneier.com/academic/archives/1997/04/the_risks_of_key_rec.html}
\end{itemize}

\textbf{In-class Video:}
\newline
\indent Dark Web Academy, ``How PGP Works,'' Mar. 26, 2016: \url{https://www.youtube.com/watch?v=CHi2RclGvIM}
\vskip.15in
\textbf{Hands-on with:}
    Protonmail, Tutanota, PGP

\vskip.25in
\noindent\textbf{Week 7 - How I Learned to \emph{Keep} Worrying} (April 12, 2017)
\vskip.07in
\textbf{Readings:}
\begin{itemize}
    \item Eben Moglen, ``Snowden and the Future,'' Dec. 4, 2013, Part IV: 
    \newline
    \url{http://www.snowdenandthefuture.info/PartIV.html}

    \item Front Line Defenders, \emph{Workbook on Security: Practical Steps for Human Rights Defenders at Risk} (2015), Chapter 5: 
    \url{https://www.frontlinedefenders.org/en/resource-publication/workbook-security-practical-steps-human-rights-defenders-risk}
    
    \item B. Schneier, K. Seidel, and S. Vijayakumar, \emph{A Worldwide Survey of Encryption Products} (2016):  \url{https://www.schneier.com/academic/archives/2016/02/a_worldwide_survey_o.html}
\end{itemize}

\textbf{Last Session:} Let's stay in touch with the tools we learned about!

%%%%%% END 
\end{document} 
